A recipe for extremely reproducible enrichment analysis (and why we need it)

### 2023-04-27 manuscript.Rmd
minor edits

This problem is not limited to bioinformatics.
In the field of drug development, according to a 2011 study by Bayer, only 25% of projects could be
replicated to a high degree @Prinz2011-ht.
When the US Biotech firm Amgen set out to reproduce 53 landmark cancer studies, they were able to
reproduce only 6 (11%) @Begley2012-di.

It is clear that as a field, the life sciences need to raise the rate of reproducibility or risk the
public perceiving it as unreliable and wasteful.

We propose the five pillars (**Figure 2**), as a framework to guarantee computational reproducibility.
However, these concepts may be difficult for novice analysts, and so we provide a step-by-step
protocol to guide them on how to customise boilerplate Dockerfile and Rmarkdown scripts to suit their
project's needs.

Although we aimed to make this process as easy as possible, it requires a substantial time outlay, which
we estimate at 15-20 hours.
This is significantly longer than using a web-based tool that provides results in just a few
minutes, but we argue the time investment is well spent.
It ensures analytical reproducibility, assists with auditing and makes the process of reproduction
simple and fast (Box 1), which should aid the early identification of errors.

@@

Although we hope that extremely reproducible research will become more widespread, we do not
anticipate any dramatic changes in researcher behaviour unless the motivating factors are augmented.
Reward structures in academia encourage individuals to publish many articles, and reproducibility
is not a significant factor unless the issues are extreme enough to provoke a retraction by the journal,
which could tarnish the reputation of an author or their institution.
Solutions to this could include mandates by funding agencies or institutions @Garijo2013-zb,
and/or badging of articles that comply with best practices @Kidwell2016-pm.

>>>>

This problem is not limited to bioinformatics.
In the field of drug development, according to a 2011 study by Bayer, only 25% of projects could be
replicated to a high degree @Prinz2011-ht.
When the US Biotech firm Amgen set out to reproduce 53 landmark cancer studies, they were able to
reproduce only 6 (11%) @Begley2012-di.
It is clear that as a field, the life sciences need to raise the rate of reproducibility or risk the
public perceiving it as unreliable and wasteful.

We propose the five pillars (**Figure 2**), as a framework to guarantee computational reproducibility.
However, these concepts may be difficult for novice analysts, and so we provide a step-by-step
protocol to guide them on how to customise template Dockerfile and Rmarkdown scripts to suit their
project's needs.
Although we aimed to make this process as easy as possible, it requires a substantial time outlay, which
we estimate at 15-20 hours.
This is significantly longer than using a web-based tool that provides results in just a few
minutes, but we argue the time investment is well spent.
It ensures analytical reproducibility, assists with auditing and makes the process of reproduction
simple and fast (Box 1), which should aid the early identification of errors.

@@

Although we hope that extremely reproducible enrichment analysis will become more widespread, we do not
anticipate any dramatic changes in researcher behaviour unless the motivating factors are augmented.
Reward structures in academia favour career progression for individuals who publish many articles
as opposed to individuals who publish fewer but more rigourously conducted articles.
Solutions to this could include mandates by funding agencies or institutions @Garijo2013-zb,
and/or badging of articles that comply with best practices @Kidwell2016-pm.
